Inferring the distribution and rate of Earth-like exoplanets \\ from noisy individual-planet inferences

### 2014-04-20 document/ms.tex
better discussion of into materials

In this \paper, we are asking the question: \emph{what constraints can we put
on the occurence rate of exoplanets given all the light curves measured by
\project{Kepler}?}
Mathematically, this involves evaluating the so-called \emph{marginalized
likelihood function}
\begin{eqnarray}\eqlabel{crazylike}
p(\{\data_k\}\,|\,\ratepars) &=&
    \int\left[\prod_{k=1}^K p(\data_k\,|\,\entry_k)\right ]
    \,p(\{\entry_k\}\,|\,\ratepars)
    \dd\{\entry_k\}
\end{eqnarray}
where $\{\data_k\}$ is the set of all light curves, \ratepars\ is a vector of
parameters describing the occurence rate function $\rate_\ratepars(\entry)$
and \entry\ is a vector of physical parameters describing a planetary system
(period, radius ratio, \etc).

It is generally computationally intractable to compute the likelihood in
\eq{crazylike} because it involves the numerical evaluation of an extremely
 high dimensional---the number of physical parameters times the number of
targets---and computationally costly integral.
Instead, we can use existing catalogs of planet candidates to approximate the
marginalization integral.
For our purposes, this is the definition of \emph{hierarchical inference}.

A very simple example of such an inference is a procedure that should be very
familiar: the histogram.
In the limit of negligible uncertainties, the maximum marginalized likelihood
for the occurence rate is a histogram of the objects in the catalog (see
\app{histogram} for the derivation).

>>>>

In this \paper, we are asking the question: \emph{what constraints can we put
on the occurence rate of exoplanets given all the light curves measured by
\kepler?}
Mathematically, this involves evaluating the so-called \emph{marginalized
likelihood function}
\begin{eqnarray}\eqlabel{crazylike}
p(\{\data_k\}\,|\,\ratepars) &=&
    \int\left[\prod_{k=1}^K p(\data_k\,|\,\entry_k)\right ]
    \,p(\{\entry_k\}\,|\,\ratepars)
    \dd\{\entry_k\}
\end{eqnarray}
where $\{\data_k\}$ is the set of all light curves, \ratepars\ is a vector of
parameters describing the occurence rate function $\rate_\ratepars(\entry)$
and \entry\ is a vector of physical parameters describing a planetary system
(period, radius ratio, \etc).

It is generally computationally intractable to compute the likelihood in
\eq{crazylike} because it involves a high dimensional numerical integration of
a function where each evaluation of the integrand touches \emph{every data
point}.
Instead, we can use existing catalogs of planet candidates to approximate the
marginalization integral.
For our purposes, this is the definition of \emph{hierarchical inference}.

A catalog is a dimensionality reduction of the data that it is based on where
all the (relevant) information has been retained (we hope!).
In the context of \kepler, the catalog reduces the set of downloaded pixel
time series (approximately XXXX for the average \kepler\ target) to
probabilistic constraints on a handful of physical parameters---we'll call
them \entry---like the orbital period and radius.
If we take this set of parameters as \emph{sufficient statistics} of the data
then we can, in theory, compute \eq{crazylike}---up to an unimportant
constant---without ever looking at the raw data again.
This is true because the catalog provides constraints on the posterior
probability of the parameters $\{\entry_k\}$ under some ``interim prior''
\interim
\begin{eqnarray}\eqlabel{crazypost}
p(\{\entry_k\}\,|\,\{\data_k\},\,\interim) &=&
\frac{p(\{\entry_k\}\,|\,\interim)}{p(\{\data_k\}\,|\,\interim)}\,
\prod_{k=1}^K p(\data_k\,|\,\entry_k) \quad.
\end{eqnarray}
We can then multiply the integrand in \eq{crazylike} by
\begin{eqnarray}
\frac{p(\{\entry_k\}\,|\,\{\data_k\},\,\interim)}
     {p(\{\entry_k\}\,|\,\{\data_k\},\,\interim)}
\end{eqnarray}
and use \eq{crazypost} to find
\begin{eqnarray}
\frac{p(\{\data_k\}\,|\,\ratepars)}{p(\{\data_k\}\,|\,\interim)} &=&
    \int
    \frac{p(\{\entry_k\}\,|\,\ratepars)}{p(\{\entry_k\}\,|\,\interim)}\,
    p(\{\entry_k\}\,|\,\{\data_k\},\,\interim)
    \dd\{\entry_k\} \quad.
\end{eqnarray}
On the right hand side of this equation, the data come in only through the
quantity $p(\{\entry_k\}\,|\,\{\data_k\},\,\interim)$ that is provided by the
catalog!

A very simple example is a procedure that will be very familiar: the histogram.
If you model the function $p(\{\entry_k\}\,|\,\ratepars)$ as a piecewise
constant rate function---where the bin heights are the parameters---and if the
uncertainties on the catalog are negligible compared to the bin widths then
the maximum marginalized solution for \ratepars\ is a histogram of the
catalog entries (see \app{histogram} for the derivation).
The case of non-negligible uncertainties is described by \citet{hogge} using a
method similar to the one discussed below.
The limiting noiseless case for an incomplete catalog also has an analytic
maximum likelihood result that we derive in \app{histogram}.
